---
title: "Customer Segmentation"
author: "Ilse Paniagua"
date: "2/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(fpp3)
library(magrittr)
```

## R Markdown

```{r}
hotels <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-11/hotels.csv')

```

```{r}
summary(hotels)
```

The dataset has one row per booking.
```{r}
#Clean up country of origin
#theory: international customers may rebook more frequently
```

Creating a date variable from year, month, and day.
```{r}
hotels %<>%  
  mutate(date = glue::glue("{arrival_date_year}-{arrival_date_month}-{arrival_date_day_of_month}"),
         date = parse_date(date, format = "%Y-%B-%d")) 
```

# City Hotels

## Bookings

Weekly seasonality is very strong. There are also some outliers in the remainder that are worth explorting. They have unusually high values at 
```{r}
#Timeseries decomposition
hotels %>% 
  filter(hotel == "City Hotel") %>%
  select(date, everything()) %>% 
  arrange(date) %>% 
  count(date) %>% 
  rename(daily_bookings = n) %>% 
  tsibble::as_tsibble() %>% 
  STL(daily_bookings ~ season(window = Inf)) %>%
  autoplot()

#Which dates have outlier values?
hotels %>% 
  filter(hotel == "City Hotel") %>%
  select(date, everything()) %>% 
  arrange(date) %>% 
  count(date) %>% 
  rename(daily_bookings = n) %>% 
  tsibble::as_tsibble() %>% 
  STL(daily_bookings ~ season(window = Inf)) %>%
  select(date, daily_bookings, remainder) %>%
  arrange(desc(remainder)) %>%
  head(15)
```

## Cancelations

How do cancelations vary over time?


Which types of customers have a higher probability of canceling?


# Resort Hotels

## Bookings

Remainder terms are lower for the resort hotel. The trend only varies by about 25 bookings over the series. 2015-12-05 was an outlier date for both the city and resport hotels.
```{r}
#Timeseries decomposition
hotels %>% 
  filter(hotel == "Resort Hotel") %>%
  select(date, everything()) %>% 
  arrange(date) %>% 
  count(date) %>% 
  rename(daily_bookings = n) %>% 
  tsibble::as_tsibble() %>% 
  STL(daily_bookings ~ season(window = Inf)) %>%
  autoplot()

#Which dates have outlier values?
hotels %>% 
  filter(hotel == "Resort Hotel") %>%
  select(date, everything()) %>% 
  arrange(date) %>% 
  count(date) %>% 
  rename(daily_bookings = n) %>% 
  tsibble::as_tsibble() %>% 
  STL(daily_bookings ~ season(window = Inf)) %>%
  select(date, daily_bookings, remainder) %>%
  arrange(desc(remainder)) %>%
  head(15)
```


